\section{Theoretical Basis and Methods}
% The theoretical basis of the XY model and the Kosterlitz-Thouless transition is rooted in the study of phase transitions and critical phenomena in two-dimensional systems~\cite{KosterlitzThouless1973}.
The XY model is a classical spin model, that generalizes the Ising model with a continuous two dimensional spin vector with unit length $\vec{s} = (s_x, s_y)$.
Due to the normilization of the spin vector, each spin can be described by only one parameter $\vec{s}_i = (\cos\theta_i, \sin\theta_i)$.
In the absence of an external magnetic field, the Hamiltonian of the XY model is thus given by:
\begin{equation}
\mathcal{H} = -J \sum_{\langle i,j \rangle} s_i \cdot s_j = -J \sum_{\langle i,j \rangle} \cos(\theta_i - \theta_j)
\end{equation}
where $J$ is the coupling constant, $\langle i,j \rangle$ denotes nearest-neighbor pairs of spins, and $\theta_i$ is the angle of the spin at site $i$.
Just like the Ising model, below $T_c$, the system exhibits quasi-long-range order, while above $T_c$, the system is in a disordered phase.
However, the correlation length is not characterized by a typical power-law anymore, but rather behaves exponentially ~\cite{Kosterlitz1974}.
In the case of two dimensions, the Kosterlitz-Thouless transition occurs in the XY model at a critical temperature $T_c \approx 0.8929J$~\cite{Hasenbusch2005}.

\subsection{Markov Chain Monte Carlo Simulation}
A Markov Chain Monte Carlo Simulation (MCMC) works by generating a sequence of configurations, where each configuration is generated from the previous one by applying a stochastic update rule.
In general, for such an algorithm to generate physically meaningful configurations, it must satisfy two key properties: Ergodicity and detailed balance.
A processs is said to be ergodic, if it can reach any possible mictrostate within the complete state space of $\Omega$, i.e. there must be a non zero probability to reach state $b$ from state $a$~\cite{PetschliesUrbach2026}.
Furthermore, a process is said to fulfill detailed balance, if, for a distribution $\pi$ and a transition matrix $P$, it fulfills the condition $\pi_\nu p_{\nu\mu} = \pi_\mu p_{\mu\nu}$, i.e. the flow between states, or the expected number of transitions per unit time, $\nu \rightarrow \mu$ must equal to its reverse $\nu \leftarrow \mu$~\cite{PetschliesUrbach2026}.
Detailed balance is a stronger condition and ensures the correct target distribution at equilibrium.

In order to give an overview of the methods used to simulate the XY model, we will briefly discuss the two main algorithms: the Metropolis algorithm and the Wolff cluster algorithm.
Both algorithms are based on the principle of MCMC simulation, allowing us to sample configurations of the system according to their Boltzmann weight~\cite{Metropolis1949}.
It can be shown that those indeed fulfill the conditions of ergodicity and detailed balance~\cite{PetschliesUrbach2026,Wolff1989}.
% The Metropolis algorithm updates the spins one at a time, while the Wolff cluster algorithm updates clusters of spins simultaneously, which can lead to faster convergence and reduced autocorrelation times, especially near critical points~\cite{Wolff1989}.
In our simulations, we will compare the results obtained from both algorithms and discuss their advantages and disadvantages in the context of simulating the XY model.

\subsection{Metropolis algorithm}
Based on the MCMC method, the Metropolis algorithm is a widely used technique for simulating statistical systems~\cite{Metropolis1949}.
In the context of the XY model, the Metropolis algorithm works as follows:

\begin{enumerate}
  \item Initialize the system with a random configuration of spins.
  \item For each Monte Carlo step, select a spin at random and propose a new angle for that spin by adding $\Delta \theta \sim U[0, 2\pi)$
  \item Calculate the change in energy $\Delta E$ resulting from the proposed update.
  \item Accept the proposed update with a probability given by the Metropolis criterion:
  \begin{equation}
    P_{\text{accept}} = \min\left(1, e^{-\Delta E / k_B T}\right)
  \end{equation}
  \item If the update is accepted, update the configuration; otherwise, keep the current configuration.
  \item Repeat steps 2-5 for a large number of Monte Carlo steps.
\end{enumerate}

The \texttt{Metropolis()} method implements single-spin flip dynamics exactly as described above.
One full sweep iterates over all $N_x \times N_y$ sites sequentially.
% For each lattice site $(x,y)$, it:
% \begin{itemize}
%   \item Computes the local energy $E_{\text{now}}$ using interactions with all four nearest neighbors
%   \item Proposes $\theta_i' = \theta_i + \delta$ where $\delta \sim U[0, 2\pi)$ via \texttt{randomAngle()}
%   \item Computes the new local energy $E_{\text{after}}$ 
%   \item Accepts the move if $E_{\text{after}} < E_{\text{now}}$ or with probability $e^{-(E_{\text{after}} - E_{\text{now}})/T}$ via \texttt{random()}
%   \item Updates $\theta_i \leftarrow \theta_i + \delta \mod 2\pi$
% \end{itemize}

\subsection{Wolff algorithm}
The Wolff algorithm is an efficient cluster update method that dramatically reduces critical slowing down near phase transitions~\cite{Wolff1989}.
Unlike single-spin flips, it builds and flips entire correlated clusters of spins:
\begin{enumerate}
  \item Chose a random reflection angle $\rho \sim U[0, 2\pi)$ and a random lattice site $s_i$ as the clusters root.
  \item Flip $s_i \rightarrow s_i = (\cos(2\rho - \theta_i + \pi), \sin(2\rho - \theta_i + \pi))$
  \item Visit all neighboring sites, and activate the bond to $s_i$ with probability
  $$
    p_{i,j} = 1 - \exp \left\{2/T \cdot \cos(\rho - \theta_i)\cos(\rho - \theta_j)\right\} 
  $$
  \item If the bond $s_i$-$s_j$ is activated, $s_j$ will be marked to be flipped next to join the cluster.
  \item Repeat steps 2-4 for all marked bonds, that are not in the cluster already.
  % \item Select a random seed spin $s_0$ and a random reflection angle .
  % \item Build a cluster by adding nearest neighbors with probability $p_{\rm add} = 1 - \exp\left[\min(0, 2\Delta\theta \cos(\phi))\right]$, where $\phi$ is the relative angle after reflection.
  % \item Flip all spins in the cluster by $\theta_i \to 2r - \theta_i \mod 2\pi$.
  % \item Repeat until a target number of spins have been flipped (here: $N_xN_y$ total).
\end{enumerate}

In the \texttt{Wolff()} method, clusters are grown using a stack-based algorithm with periodic boundary conditions, to keep track of the marked bonds.
One full sweep will attempt to flip clusters, as long as the expected cluster size of the next cluster would not exceed a total number of flipped spins of $N_x \times N_y$.
% making it particularly effective for studying the XY model's BKT transition.

\subsection{Observables}
In order to analyze the behavior of the system in response to the temperature, some obervables need to be measured.
These observables are known to exhibit interesting properties near the phase transition.

The magnetization can be thought of as the average alignment of the spins in the system of size $L \times L$, and is given for a specific spin configuration $k$ by
$$
  M_k = \frac{1}{L^2}\sqrt{\left(\sum_{i=0}^{L^2}\cos \theta_i\right) + \left(\sum_{i=0}^{L^2}\sin \theta_i\right)} \;.
$$
This can then be used to get the average magnetization over $N$ total spin configurations
$$
  \langle M \rangle = \frac{1}{N} \sum_{i=0}^{N} M_i \;.
$$
Another interesting property is the magnetic susceptibility $\chi$, which quantifies the response of the system to a applied external magnetic field.
It is expected to diverge near the critical temperature~\cite{Kosterlitz1974}, and will thus be an excellent observble to analyze the system near the critical temperature.
Intuitively, it can be thought of as the variance of the magnetization, and is given by
$$
  \chi = \frac{L^2}{T} ( \langle M^2 \rangle - \langle M \rangle^2) \;.
$$
Analogously to the magnetization, the energy and the specific heat of the system can be defined as
\begin{align}
  E_k =& -\frac{1}{L^2} \sum_{\langle i,j \rangle} \cos(\theta_i - \theta_j) \;,\\
  C =& \frac{L^2}{T^2} ( \langle E^2 \rangle - \langle E \rangle^2) \;.
\end{align}

\subsection{Critical Exponents}
Near a phase transition, the behavior of physical quantities can be described using critical exponents.
In the context of the XY Model, these physical quantities diverge even faster than the general power law behavior~\cite{Kosterlitz1974}.
However, by expressing them in terms of the correlation length $\xi$, which scales exponentially $\xi \sim \exp(b\cdot t^{-1/2})$, the usual shape of the critical exponents can be recovered, where $t = \frac{T_c - T}{T}$ is the reduced temperature.
Near $T_c$, the susceptibility and specific heat scale according to
\begin{align}
  \chi \sim & \xi^{\tilde{\gamma}} \hspace{1cm}, \tilde{\gamma} = \frac{\gamma}{\nu} = 2 - \eta\\
  C_v \sim & \xi^{\tilde{\alpha}} \hspace{1cm}, \tilde{\alpha} = \frac{\alpha}{\nu}
\end{align}
where $\nu$ is the critical exponent of the correlation length $\xi$, $\gamma$ that for the susceptibility $\chi$, $\alpha$ that for the specific heat $C_v$.
A value of $\eta = 1/4$ is the same as that of the Ising model predicted by Onsager.~\cite{Kosterlitz1974}

\subsection{Finite Size Scaling (FSS)}
Finite size scaling analyzes how the observables behave as a function of system size $L = \sqrt{N}$ near criticality ~\cite{Privman1989}.
In theory, the correlation length diverges at $T_c$, but for a finite system size it is bounded by $L$.
Thus, one expects the observables to exhibit such a scaling behavior.
So, for the XY model near the phase transition, the susceptibility $\chi_L(T)$ scales as

\begin{equation}
  \chi(L,t) \approx L^{\gamma/\nu} \,\hat{\chi}\!\left( L \exp\!\left(-b\, t^{-\nu}\right) \right) \;,
  \label{eq:Susceptibility_Scale}
\end{equation}
with $t$ the reduced temperature, $b$ a non-universal constant, and $\nu$ is a critical exponent related to the correlation length~\cite{Sale2021}.
By scaling the data for various lattice sizes accordingly, the curves will collapse onto a single universal curve, allowing the exctraction of the critical exponents and critical temperature.

Another method to extract these quantities, is by fitting a line through the peak positions $T_c(L)$ of the susceptibility as a function of lattice size.
The peak positions scale according to 
\begin{equation}
  T_c(L) - T_c(\infty) \propto \frac{1}{\log (L)^{1/\nu}} \;,
  \label{eq:bkt_scaling}
\end{equation}
so $T_c$ and $\nu$ can be obtained from the fit parameters~\cite{Sale2021}.

\subsection{Autocorrelation}
Generally speaking, the autocorrelation function $C_X(t)$ of an observable $X$ measures how correlated the values of $X$ are at different time steps $t$.
It is defined as:
\begin{equation}
  C_X(t)\;\stackrel{\mathrm{def}}{=}\;\big\langle (X_i-\mu)\,(X_{i+t}-\mu)\big\rangle
\end{equation}
where $\mu$ is the mean of $X$ and the average $\langle \cdot \rangle$ is taken over sufficiently large sample sizes.
The normalized autocorrelation function $\Gamma_X(t)$ is then given by
\begin{equation}
  \Gamma_X(t)\;\stackrel{\mathrm{def}}{=}\;\frac{C_X(t)}{C_X(0)} \;,
  \label{eq:Autocorrelation_Function}
\end{equation}
and is a measure of how quickly the system "forgets" its past configurations.
$C_X(t)$ depends only on the time difference $t$, and typically decays exponentially,
\begin{equation}
  C_X(t)\propto \sum_{n\ge 0} A_n\,\exp\!\left(-\frac{|t|}{\tau_n}\right),
  \qquad \tau_0>\tau_1>\cdots \in \mathbb{R}^+ .
\end{equation}
For sufficiently large $|t|$ only the largest correlation time $\tau_0$ survives.
Hence, one defines the exponential autocorrelation time by
\begin{equation}
  \tau_{\exp}\stackrel{\mathrm{def}}{=}\limsup_{t\to\infty}\frac{-t}{\ln\!\bigl|\Gamma_X(t)\bigr|}\,.
  \label{eq:Autocorrelation_Time}
\end{equation}
It is the time needed for the autocorrelation to be reduced by a factor of order $e$, and needs to be taken into account when generating statistically independen MCMC configurations.~\cite{PetschliesUrbach2026}